Probing heavy ion collisions using quark and gluon jet substructure

### 2018-02-27 qg_ML.tex
YT updated the draft.

\abstract{
We study the phenomenon of jet quenching utilizing quark and gluon jet substructures as independent probes of heavy ion collisions. We exploit jet and subjet features to highlight differences between quark and gluon jets in vacuum and in a medium with the jet-quenching model as implemented implemented in \textsc{Jewel}. We begin with a physics-motivated, multivariate analysis of jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity. In comparison, we employ state-of-the-art image-recognition techniques by training a deep convolutional neutral network on jet images. To systematically extract jet substructure information, we introduce the telescoping deconstruction framework exploiting subjet kinematics at multiple angular scales. We draw connections to the soft-drop subjet distribution and illuminate medium-induced jet modifications using Lund diagrams. We find that the quark gluon discrimination performance worsens in heavy ion jets due to significant soft event activity affecting the soft jet substructure. Our work suggests a systematically improvable framework for studying modifications to quark and gluon jet substructures and to facilitate direct comparisons between analytic calculations, simulations and measurements in heavy ion collisions.
}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Comp_delta_m2.pdf}
	   \caption{Distributions of $\delta m=\sqrt{m_{Jet}^2-m_{groomed-Jet}^2}$ for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
\label{fig:comp_delta_m2}
\end{figure}

Using the general idea of telescoping jet substructure \cite{Chien:2017xrb} we also look at the difference between the square of jet mass for ungroomed and soft-drop groomed jets: $\delta m \equiv \sqrt{m_{Jet}^2-m_{groomed-Jet}^2}$. Since soft-drop removes soft, wide angle radiation from jets, $\delta m$ disentangles hard, collinear radiation and probes the soft radiation that is dropped. Figure:~\ref{fig:comp_delta_m2} shows the $\delta m$ distributions for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions simulated in \textsc{Jewel}. We see that the observable $\delta m$ in pp collisions exhibits differences between quark jets and gluon jets due to the asymmetric impact of softdrop on gluon jet mass and structure. On the other hand, the distributions for quark and gluon jets become very similar in PbPb collisions, albeit shifted towards higher values, and consequently a broader distribution as compared to pp collisions. This hints that soft event activities in \textsc{Jewel} can end up smearing the differences between quark and gluon jets, which leads to the lower classification performance as we will show in the next section. Another interesting feature highlighted by the $\delta m$ distributions is the peak at zero, i.e. the fraction of jets who's mass is essentially unaffected by grooming. As shown in Fig:~\ref{fig:comp_delta_m2}, the distribution at x-axis $\approx 0$ is higher for gluon jets as opposed to quark jets and this trend persists even after quenching.

@@

\section*{Acknowledgments}
The authors are grateful to Patrick Komiske and Eric Metodiev for collaborating on the development of telescoping deconstruction and contributions to the Appendix. The authors also thank Fr\'ed\'eric Dreyer, Benjamin Nachman, Joern Putschke, Jesse Thaler, Konrad Tywoniuk and Sevil Salur for helpful discussions. We thank Gavin Salam and the participants of the CERN TH institute ``{\sl Novel tools and observables for jet physics in heavy ion collisions}" and the 5-th Heavy Ion Jet Workshop for bringing to our attention the Lund Diagram and its usage in studying jet modifications. RKE also thanks the organizers of the DS@HEP 2017 workshop at FermiLab for an excellent introduction on machine learning techniques. YTC also thanks the organizers of the BOOST2015 conference where the idea of telescoping deconstruction was first presented. YTC was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008. RKE acknowledges supports in part by the U.S. Department of Energy Office of Science, Office of Nuclear Physics under Award Number DE-FG02-92ER-40713.

@@

In this Appendix, we present the framework of telescoping deconstruction to systematically probe aspects of jet formation via its fragmentation basis. By scanning around dominant energy flows with multiple angular resolutions \cite{Chien:2013kca,Chien:2014hla}, one can efficiently quantify the radiation pattern. We demonstrated the use of the framework in the main text for quark/gluon discrimination and to study the medium modification in heavy ion collisions and in this appendix, we take the approach to a much different problem at the LHC: identifying and tagging boosted quark/gluon jets,  boosted $W$ and tops at the LHC. Crucially, the framework involves a fixed-order organization of individual observables which allows systematically improvable jet studies. Explicit examples of the $W$ isolation \cite{Chien:2017xrb} and exposing the $W$ boson in a top jet are presented. This highlights the physically meaningful nature of each TD observables and demonstrates their collective power as a representation.

@@

We aims to encapsulate relevant physics information in simple, physical observables derived with the TD framework. These variables follow the perturbative expansion of QCD and the parton shower picture and therefore encapsulate perturbative and non-perturbative physics information. Similar to fixed-order perturbative expansions and parton shower splitting kernels \cite{Nagy:2017ggp}, the expansion is ordered by the number $N$ of exclusively reconstructed subjets. At each order, $N$ axes are determined by finding the dominant energy flow directions in the rapidity-azimuth plane~\cite{Stewart:2010tn,Chien:2013kca,Stewart:2015waa,Thaler:2015xaa}, which is partitioned into energy flow regions determined by the nearest axis. Jets at multiple angular resolutions are probed simply by the kinematics of subjets consisting of particles within different distances $R_T$ from the energy flow axes.

\begin{figure}[t]
\centering
\includegraphics[width=.32\columnwidth]{figures/topjetT1_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT2_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT3_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT1_masses2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT2_masses2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT3_masses2.pdf}
\caption{\label{fig:tjet}(Top row) The TD of a top jet at T1 (left panel), T2 (middle panel) and T3 (right panel) orders. The crosses are the Winner-Take-All $k_T$ axes. The straight lines are the exclusive subjet boundaries. Particles are sized according to their transverse momenta. (Bottom row) The subjet masses at T1 (left panel), T2 (middle panel), and T3 (right panel) orders as the telescoping radius $R_T$ is varied, corresponding to the region of the same color in the top row. The presence of the $W$ in the top jet can be clearly seen at T1 and T2 orders by the plateau of the red line beginning around $R_T=0.3$.}
\end{figure}

@@

The TD observables fall into two categories~\cite{Chien:2017xrb}: the {\sl subjet topology}, which is described by the axes and subjet transverse momenta, and the {\sl subjet substructure}, quantified by the subjet masses~\footnote{{\sl Subjet charge}~\cite{Krohn:2012fg} information can be included in this framework as well}. As the telescoping subjet order $N$ (T$N$ order) increases, more jet energy is covered by the subjets and the number of subjet radii sampled can be systematically decreased. In the large $N$ limit, the subjets reduce to individual particles with the full jet information. The telescoping deconstruction allows one to exploit features both within each subjet and among all the subjets. See Fig:~\ref{fig:tjet} for the TD of a top jet for $N = 1$, $2$, and $3$ where the $W$ mass resonance can clearly be seen in the bottom panels.

@@

Since $W$ and top tagging are mass resonance searches where the jet mass is the most natural and powerful discriminating variable, we disentangle mass information to probe how much additional information can be exploited for tagging. Information from the hard process about the overall jet kinematics is eliminated by translating each jet to a frame where $y$ and $\phi$ of the jet are both zero. For $W$ and top tagging, $W$ jets were generated by $pp\to WW (\to \text{hadrons})$ and top jets were generated by $pp\to t\bar t (\to \text{hadrons})$, with background jets taken from QCD dijets. Final-state non-neutrino particles are clustered into $R=1.0$ jets with rapidity $|y|<1.5$. Signal and background jets identically populate a five-bin $1/p_T^4$ histogram in transverse momentum and a three-bin uniform histogram in mass between 75 GeV and 85 GeV for $W$ tagging and 160 GeV to 180 GeV for top tagging. Telescoping subjets are constructed with $R_T$ in steps of 0.05 between 0.05 and 0.4 in quark/gluon discrimination, and between 0.05 to 0.95 in $W$ and top tagging. Reducing the number of radius sampling is left for future work. For each problem, 200k events are generated for signal and background, with 10\% used for validation, 15\% for testing, and the remaining 75\% for training.

A neural network consisting of three dense layers with 100 nodes each is trained on the telescoping deconstruction observables up to T$N$ order. The training can be performed on a typical laptop CPU in fewer than five minutes. One could also use boosted decision trees to combine variables \cite{Chien:2017xrb}. We compare the performance of our method to jet images using a CNN as done in the main text. The jet images are size $33\times 33$ and span a $2R\times 2R$ patch of the rapidity-azimuth plane with the intensity of each pixel corresponding to the total $p_T$ of particles in the pixel. The jet images are pre-processed and standardized according to the procedure in \Ref{Komiske:2016rsd}. The CNN architecture consists of three 48-filter convolutional layers with filter sizes of $8\times 8$, $4\times 4$, and $4\times 4$ followed by a 128-unit dense layer and a 2-unit softmaxed output layer. A $2\times 2$ maxpooling is performed after each convolutional layer with a stride length of 2. The dropout rate was taken to be 0.1 for all layers. All neural networks are implemented using the Python deep learning library Keras~\cite{keras} with the Theano backend~\cite{bergstra2010theano}. Rectified linear unit (ReLU) activation functions~\cite{nair2010rectified} and He-uniform model weight initialization~\cite{heuniform} are used. The networks are trained using the Adam algorithm~\cite{adam} with a learning rate of $10^{-3}$ and a batch size of 256 for 50 epochs with a patience parameter of 8, and the best model is selected based on validation set performance.

The performance of the trained models can be captured in a slightly different ROC curve (compared to the main text) where we plot the inverse of the background mistag rate at different signal efficiencies, thus a higher curve indicates better classification performance. \Fig{ROCs} shows the ROC curves for the three tagging problems of the models trained on the cumulant telescoping deconstruction observables up to T$N$ order for $N\in\{1,2,3,4\}$ and the CNNs trained on jet images. The T$N$ performance converges quickly and is comparable to the performance of the jet images approach. The CNN architecture has not been tuned exhaustively, therefore its ROC curves serve to give a general sense of performance.

\begin{figure}[h]
\centering
\includegraphics[width=.32\columnwidth]{figures/T1HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T2HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T3HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T1HeavyMass_qcd.pdf}
\includegraphics[width=.32\columnwidth]{figures/T2HeavyMass_qcd.pdf}
\includegraphics[width=.32\columnwidth]{figures/T3HeavyMass_qcd.pdf}
\caption{\label{masses}The heavy subjet mass distributions at T1 (left panels), T2 (middle panels), and T3 (right panels) orders for multiple subjet radii $R_T$ of top jets (top row) and QCD background jets (bottom row). The presence of the $W$ in top jets is evident by the peaks at the $W$ mass at T1 and T2 orders. At T3 order, the heavy subjet mass distributions are QCD-like and are narrower (more quark-like) in top subjets than the wider (more gluon-like) QCD subjets. These features highlight the ability of the telescoping deconstruction to probe subjet substructure.}
\end{figure}

@@

In addition to being a useful jet representation, the TD allows physical information to be easily extracted. \Fig{masses} shows the heavy subjet mass \cite{Chien:2010kc} distributions at T1, T2, and T3 orders, scanned over different telescoping radii $R_T$, for top jets and their QCD background. As $R_T$ is increased, the top jet T1 subjet mass distributions transition from QCD-like, to peaked at the $W$ mass, to peaked at the top mass. In contrast, the background QCD jets do not peak at the $W$ mass for any $R_T$ and transition from QCD-like to more top-like as they acquire mass at larger radii. The top jets at T2 order clearly and automatically show the $W$ peak which is completely absent in the background distributions. At T3 order, both top and background distributions appear QCD-like, with the wider background distributions due to the prevalence of gluon subjets. Thus we find that TD extracts the relevant and significant subjet substructure observables in order to efficiently perform at its task, i.e. quark jet vs gluon jet or boosted W/tops vs QCD jets.

>>>>

\abstract{
We study the phenomenon of jet quenching utilizing quark and gluon jet substructures as independent probes of heavy ion collisions. We exploit jet and subjet features to highlight differences between quark and gluon jets in vacuum and in a medium with the jet-quenching model implemented in \textsc{Jewel}. We begin with a physics-motivated, multivariate analysis of jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity. In comparison, we employ state-of-the-art image-recognition techniques by training a deep convolutional neutral network on jet images. To systematically extract jet substructure information, we introduce the telescoping deconstruction framework exploiting subjet kinematics at multiple angular scales. We draw connections to the soft-drop subjet distribution and illuminate medium-induced jet modifications using Lund diagrams. We find that the quark gluon discrimination performance worsens in heavy ion jets due to significant soft event activity affecting the soft jet substructure. Our work suggests a systematically improvable framework for studying modifications to quark and gluon jet substructures and to facilitate direct comparisons between theoretical calculations, simulations and measurements in heavy ion collisions.
}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Comp_delta_m2.pdf}
	   \caption{Distributions of $\delta m=\sqrt{m_{\rm jet}^2-m_{\rm groomed-jet}^2}$ for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
\label{fig:comp_delta_m2}
\end{figure}

Using the general idea of telescoping jet substructure and variability \cite{Chien:2017xrb} we also look at the difference between the square of jet mass for ungroomed and soft-drop groomed jets: $\delta m \equiv \sqrt{m_{\rm jet}^2-m_{\rm groomed-jet}^2}$. Since soft-drop removes soft, wide angle radiation from jets, $\delta m$ disentangles hard, collinear radiation and probes the soft radiation that is dropped. Figure~\ref{fig:comp_delta_m2} shows the $\delta m$ distributions for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions simulated in \textsc{Jewel}. We see that the observable $\delta m$ in pp collisions exhibits differences between quark jets and gluon jets. %due to the asymmetric impact of softdrop on gluon jet mass and structure.
On the other hand, the distributions for quark and gluon jets become very similar in PbPb collisions, albeit shifted towards higher values, and consequently become broader as compared to pp collisions. This hints that soft event activities in \textsc{Jewel} can end up smearing the differences between quark and gluon jets, which leads to the lower classification performance as we will show in the next section. Another interesting feature highlighted by the $\delta m$ distributions is the peak at zero, i.e. there is significant fraction of jets of which the masses are essentially unaffected by soft-drop. In other words, for such jets no particles are removed by soft-drop. As shown in Figure~\ref{fig:comp_delta_m2}, the distribution at $\delta m \approx 0$ is higher for gluon jets as opposed to quark jets. This trend persists even after quenching.

@@

\section*{Acknowledgments}
The authors are grateful to Patrick Komiske and Eric Metodiev for collaborating on the development of telescoping deconstruction and contributions to the Appendix. The authors also thank Fr\'ed\'eric Dreyer, Benjamin Nachman, Joern Putschke, Jesse Thaler, Konrad Tywoniuk and Sevil Salur for helpful discussions. We thank Gavin Salam and the participants of the CERN TH institute ``{\sl Novel tools and observables for jet physics in heavy ion collisions}" and the 5-th Heavy Ion Jet Workshop for bringing to our attention the Lund Diagram and its usage in studying jet modifications. RKE also thanks the organizers of the DS@HEP 2017 workshop at FermiLab for an excellent introduction on machine learning techniques. YTC also thanks the organizers of the BOOST2015 conference where the ideas of telescoping deconstruction and $\delta m$ were first presented. YTC was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008. RKE acknowledges supports in part by the U.S. Department of Energy Office of Science, Office of Nuclear Physics under Award Number DE-FG02-92ER-40713.

@@

In this Appendix, we present the framework of telescoping deconstruction (TD) to systematically probe aspects of jet formation via its fragmentation basis. By scanning around dominant energy flows with multiple angular resolutions \cite{Chien:2013kca,Chien:2014hla}, one can efficiently quantify the radiation pattern. We demonstrated the use of the framework in the main text for quark/gluon discrimination and the study of medium modifications in heavy ion collisions. Here, to show the generality of TD we apply the method to different problems at the LHC: identifying and tagging boosted quark/gluon jets,  boosted $W$ and tops at the LHC. Crucially, the framework involves a fixed-order organization of individual observables which allows systematically improvable jet studies. Explicit examples of the $W$ isolation \cite{Chien:2017xrb} and exposing the $W$ boson in a top jet are presented. This highlights the physically meaningful nature of each TD observables and demonstrates their collective power as a representation.

@@

We aims to encapsulate relevant physics information in simple, physical observables derived with the TD framework. These variables follow the perturbative expansion of QCD and the parton shower picture and therefore encode perturbative and non-perturbative physics information. Similar to fixed-order perturbative expansions and parton shower splitting kernels \cite{Nagy:2017ggp}, the expansion is ordered by the number $N$ of exclusively reconstructed subjets. At each order, $N$ axes are determined by finding the dominant energy flow directions in the rapidity-azimuth plane~\cite{Stewart:2010tn,Chien:2013kca,Stewart:2015waa,Thaler:2015xaa}, which is partitioned into energy flow regions determined by the nearest axis. Jets at multiple angular resolutions are probed simply by the kinematics of subjets consisting of particles within different distances $R_T$ from the energy flow axes.

\begin{figure}[t]
\centering
\includegraphics[width=.32\columnwidth]{figures/topjetT1_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT2_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT3_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT1_masses2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT2_masses2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT3_masses2.pdf}
\caption{\label{fig:tjet}(Top row) The TD of a top jet at T1 (left panel), T2 (middle panel) and T3 (right panel) orders. The crosses are the Winner-Take-All $k_t$ axes. The straight lines are the exclusive subjet boundaries. Particles are sized according to their transverse momenta. (Bottom row) The subjet masses at T1 (left panel), T2 (middle panel), and T3 (right panel) orders as the telescoping radius $R_T$ is varied, corresponding to the region of the same color in the top row. The presence of the $W$ in the top jet can be clearly seen at T1 and T2 orders by the plateau of the red line beginning around $R_T=0.3$.}
\end{figure}

@@

The TD observables fall into two categories~\cite{Chien:2017xrb}: the {\sl subjet topology}, which is described by the axes and subjet transverse momenta, and the {\sl subjet substructure}, quantified by the subjet masses~\footnote{{\sl Subjet charge}~\cite{Krohn:2012fg} information can be included in this framework as well}. As the telescoping subjet order $N$ (T$N$ order) increases, more jet energy is covered by the subjets and the number of subjet radii sampled can be systematically decreased. In the large $N$ limit, the subjets reduce to individual particles with the full jet information. The telescoping deconstruction allows one to exploit features both within each subjet and among all the subjets. See Figure~\ref{fig:tjet} for the TD of a top jet for $N = 1$, $2$, and $3$ where the $W$ mass resonance can clearly be seen in the bottom panels.

@@

Since $W$ and top tagging are mass resonance searches where the jet mass is the most natural and powerful discriminating variable, we disentangle mass information to probe how much additional information can be exploited for tagging. Information from the hard process about the overall jet kinematics is eliminated by translating each jet to a frame where $y$ and $\phi$ of the jet are both zero. For $W$ and top tagging, $W$ jets were generated by $pp\to WW (\to \text{hadrons})$ and top jets were generated by $pp\to t\bar t (\to \text{hadrons})$, with background jets taken from QCD dijets. Final-state non-neutrino particles are clustered into $R=1.0$ jets with rapidity $|y|<1.5$. Signal and background jets identically populate a five-bin $1/p_T^4$ histogram in transverse momentum and a three-bin uniform histogram in mass between 75 GeV and 85 GeV for $W$ tagging and 160 GeV to 180 GeV for top tagging. Telescoping subjets are constructed with $R_T$ in steps of 0.05 between 0.05 and 0.4 in quark/gluon discrimination, and between 0.05 to 0.95 in $W$ and top tagging. %Reducing the number of radius sampling is left for future work. 
For each problem, 200k events are generated for signal and background, with 10\% used for validation, 15\% for testing, and the remaining 75\% for training.

A neural network consisting of three dense layers with 100 nodes each is trained on the telescoping deconstruction observables up to T$N$ order. The training can be performed on a typical laptop CPU in fewer than five minutes. One could also use boosted decision trees (BDT) to combine variables \cite{Chien:2017xrb}. We compare the performance of our method to jet images using a CNN as done in the main text. The jet images are size $33\times 33$ and span a $2R\times 2R$ patch of the rapidity-azimuth plane with the intensity of each pixel corresponding to the total $p_T$ of particles in the pixel. The jet images are pre-processed and standardized according to the procedure in \Ref{Komiske:2016rsd}. The CNN architecture consists of three 48-filter convolutional layers with filter sizes of $8\times 8$, $4\times 4$, and $4\times 4$ followed by a 128-unit dense layer and a 2-unit softmaxed output layer. A $2\times 2$ maxpooling is performed after each convolutional layer with a stride length of 2. The dropout rate was taken to be 0.1 for all layers. All neural networks are implemented using the Python deep learning library Keras~\cite{keras} with the Theano backend~\cite{bergstra2010theano}. Rectified linear unit (ReLU) activation functions~\cite{nair2010rectified} and He-uniform model weight initialization~\cite{heuniform} are used. The networks are trained using the Adam algorithm~\cite{adam} with a learning rate of $10^{-3}$ and a batch size of 256 for 50 epochs with a patience parameter of 8, and the best model is selected based on validation set performance.

\begin{figure}[t]
\centering
\includegraphics[width=.32\columnwidth]{figures/T1HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T2HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T3HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T1HeavyMass_qcd.pdf}
\includegraphics[width=.32\columnwidth]{figures/T2HeavyMass_qcd.pdf}
\includegraphics[width=.32\columnwidth]{figures/T3HeavyMass_qcd.pdf}
\caption{\label{masses}The heavy subjet mass distributions at T1 (left panels), T2 (middle panels), and T3 (right panels) orders for multiple subjet radii $R_T$ of top jets (top row) and QCD background jets (bottom row). The presence of the $W$ in top jets is evident by the peaks at the $W$ mass at T1 and T2 orders. At T3 order, the heavy subjet mass distributions are QCD-like and are narrower (more quark-like) in top subjets than the wider (more gluon-like) QCD subjets. These features highlight the ability of the telescoping deconstruction to probe subjet substructure.}
\end{figure}

The performance of the trained models can be captured in a slightly different ROC curve (compared to the main text) where we plot the inverse of the background mistag rate at different signal efficiencies, thus a higher curve indicates better classification performance. Figure~\ref{ROCs} shows the ROC curves for the three tagging problems of the models trained on the cumulant telescoping deconstruction observables up to T$N$ order for $N\in\{1,2,3,4\}$ and the CNNs trained on jet images. The T$N$ performance converges quickly and is comparable to the performance of the jet images approach. The CNN architecture has not been tuned exhaustively, therefore its ROC curves serve to give a general sense of performance.

@@

In addition to being a useful jet representation, the TD allows physical information to be easily extracted. Figure~\ref{masses} shows the heavy subjet mass \cite{Chien:2010kc} distributions at T1, T2, and T3 orders, scanned over different telescoping radii $R_T$, for top jets and their QCD background. As $R_T$ is increased, the top jet T1 subjet mass distributions transition from QCD-like, to peaked at the $W$ mass, to peaked at the top mass. In contrast, the background QCD jets do not peak at the $W$ mass for any $R_T$ and transition from QCD-like to more top-like as they acquire mass at larger radii. The top jets at T2 order clearly and automatically show the $W$ peak which is completely absent in the background distributions. At T3 order, both top and background distributions appear QCD-like, with the wider background distributions due to the prevalence of gluon subjets. Thus we find that TD extracts relevant and significant subjet substructure information in order to efficiently perform at its task, i.e. quark jet v.s. gluon jet or boosted $W$/top jets v.s. QCD jets.
